---
title: "EPP-ASM"
subtitle: "sexual debut, age-mixing, and risk group mixing"
author: ""
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    toc_float: true
    theme: "cerulean"
    keep_md: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  comment = "#>"
)
# devtool::install_github("kklot/eppasm@mixing")
library(eppasm)
# devtools::load_all()
```

This vignette documents the implementation of age- and risk-group-mixing
(mixing for short) model (Tim's) into current EPP-ASM code. The code lives at
`mixing` branch. The differences compare to the old version are:

<!-- and generating outputs to provide as Spectrum inputs. -->

- We need a new list named `mx` (perhaps put into `fp` later) containing all
  mixing parameters; `mx` enters `simmod` as the function's argument.
- Call `simmod(fp, isMixing=TRUE, mx)` to specify whether to run mixing model,
  otherwise the old model does.
- Adjustments are made to allow more risk groups (e.g., sex indexing,
  distributing of population into groups)
- The calculation of the force of infection (FOI), consequently the number of
  new infection in each time step.

## The model

The force of infection (FOI) is defined as in [Hallett, T. B., Gregson, S., et
al. (2007)](http://doi.org/10.1136/sti.2006.023606). Briefly, the FOI
($\lambda$) of a population of sex $S=\{1,2\}$, age $A=1,\dots$, and risk
group $G = 1\dots$ is calculated at each time steps as
\[\lambda_{SAG} = C_{SAG}\sum_g\sum_a C_{SAGsag} P_{sag} (1-K_{Aa}),\]
where $s,a,g$ denote the sex, age, and risk group of the opposite sex,
respectively. $C_{SAG}$ is the total number of contacts, $C_{SAGsag}$ is the
mixing pattern with other age and group, $P_{sag}$ is the tranmission risk,
and $K_{Aa}$ is the probability of correct and consistent condom usage given
the pair. All the model's terms are time-dependent but $K_{Aa}$ might not,
depending on the availability of data.

$C_{sag}$ is calculated as
\[C_{SAG} = \dfrac{M}{\tau^{\sum m\gamma_m}}\tau^m, \qquad m=G-1, \]
where $M$ is the geometric mean of the total number of contacts, $\tau$ is
common ratio between risk groups and $\gamma_m$ is the proportion of
population in the group $G$.

$C_{SAGsag}$ is calculated as
\[C_{SAGsag} = \left[ (1-\epsilon)\delta_{Gg} + \epsilon\left(\dfrac{C_{sag}N_{sag}}{\sum_g N_{sag} C_{sag}}\right)\right]\Delta_{SAsa},\]
where $\epsilon \in [0,1]$ defines whether contacts are assortative or random
with $\delta_{Gg}$ is the Koronecker delta. $N_{sag}$ is the size of the
corresponding population. $\Delta_{SAsa}$ is the age-mixing pattern of the
pair, defining the probability that a partnership is formed between an aged
$A$ subject and aged $a$ subject.

$\Delta_{Aa}$ can be defined similarly to the group mixing term with another
$\varepsilon$ defining the age-mixing pattern. Here $\Delta_{Aa}$ is defined
by fitting a log-logistic model ($f$) to the sexual partner age difference
data. For female aged $A$ and male aged $a$

\[\Delta_{Aa} = 
\begin{cases}
  f(a -A),& \qquad a \geq A\\
  0,& \text{otherwise},
\end{cases}
\]
and for male $\Delta_{Aa} = \Delta_{aA}$. This is adjusted such that
$\sum_{a=a_L}^{a=a_U}\Delta_{Aa} = 1,$ where $a_L, a_U$ denote the age at
first sex and age at last sex, respectively.

Lastly, $C_{SAGsag}$ needs to be adjusted to ensure the balance of the number
of partnership calculated from female and male, i.e.,

\[N_{SAG} C_{SAGsag} = N_{sag} C_{sagSAG}.\]

In particular, we calculate 
\[W = \dfrac{N_{SAG} C_{SAGsag}}{N_{sag} C_{sagSAG}}\]
and adjust
\begin{align}
  C_{SAGsag} &= C_{SAGsag} \sqrt{W} \newline
  C_{sagSAG} &= C_{sagSAG} / \sqrt{W}
\end{align}

## Assumptions and parameters

The following assumptions and parameters are used for prototyping the model.

### Age mixing

Age- and sex-specific mixing matrix ($\Delta_{Aa}$) follows a three parameters
log-logistic distribution ($\kappa, \rho, r$) in which the scale parameter `r`
is shifted by the age difference between male and female. 


```{r}
mx <- list(kappa = 3.04, # log-logistic shape
           rho = 2.85, # log-logistic scale
           r = 0.11) # log-logistic support
```

This distribution seems too strict. Considering options that allows female
with younger male and higher probability for the same age.

```{r echo=FALSE, fig.cap = "Fig: Age mixing"} 
log.logistic <- function(dis, ka = mx$kappa, rho = mx$rho, r = mx$r)  (ka*rho^ka)*((dis+r)^(ka-1)) / (1 + ((dis+r)^ka)^2)
out <- sapply(seq(-10, 10, .1), function(x) log.logistic(x, 3.04, 2.85, .11))
par(mar=c(5,4,1,1))
plot(out/max(out, na.rm=TRUE), type='l', ylab = "Prob.", xlab = "Age difference", xaxt='n')
lines(rev(out)/max(out, na.rm=TRUE), lty=3)
axis(1, at = seq(0, 200, by=10), label=-10:10)
legend('topleft', lty=c(1,3), legend = c("Female", "Male"))
```

It might be easier to work with the normal age-mixing formula with only one
parameter, but need data to fit.

### Risk groups mixing

The proportion in low and high risk group are assumed 80% and 20%. This is
used to distribute:

- the initial population
- the entrants, healthy and H+
- immigrants

```{r}
mx$gamma <- c(.8, .2) # Default order: Low High, sum to 1
```

If `mx$gamma` is of length one, then it is set to 1 and one risk group model
is run.

The following transition rates are the same among groups:

- Natural survival rate
- cd4_mort, cd4_initdist, cd4_prog, paedsurv_cd4dist, 
- art_mort, paedsurv_artcd4dist

Higher risk group have tripple the frequency of contacts of the immediate lower risk group.

```{r}
mx$tau <- 3 # common ratio between risk groups
```

where the number of contacts is given with an age- and sex-specific geometric mean number of contacts.

```{r}
mx$M <- matrix(c(c(1.6, 1.6, 1.6, 1.7, 1.6, 1),
                 c(1.2, 1.2, 1.1, 1.5, 1.1, 1),
                 c(15, 20, 25, 30, 40, 50),
                 c(19, 24, 29, 39, 49, 80)), 6, 4, dimnames=list(c("1529", "2024", "2529", "3039", "4049", "50+"), c("M", "F", "lo", "up")))
# expanding the age group provided in M
mx$M. <- function(x) apply(x[, 1:2], 2, rep, times = apply(x[, 3:4], 1, diff) + 1)
mx$Mx <- mx$M.(mx$M)
```

```{r echo = FALSE, results = 'asis'}
knitr::kable(mx$M, caption="Table: Risk group mixing")
```

### Condom usage
The probability of correct and consistent use of among age-groups and sexes are given as inputs. Below is a condom use data for only ±25 years old. This is age-dependent not risk group dependent which probably should be modified. 

```{r}
# prob of correct and consistent condom use by <25 and 25+
mx$chi <- matrix(c(.32, .1, .39, .07), 2, 2,
                 dimnames=list(c("M<25", "M25+"), c("F<25", "F25+")))
```

```{r echo = FALSE, results = 'asis'}
knitr::kable(mx$chi, caption="Table: Condom usage")
```

### Sexual debut

Population under the age of first sex and above the age of last sex would not have any sexual contacts in this model. For now the code covers only the case that both sexes have the same sexual debut period.

```{r}
mx$A1 = 16 # age at first sex
mx$A2 = 55 # age at last sex
```

## EPP model with single set of parameter inputs

Prepare model parameters from parameter vector input
```{r}
# devtool::install_github("kklot/eppasm@mixing")
library(eppasm)
devtools::load_all()
theta_ur <- c(-0.63758, -2.76655, -1.26204, 1996.65945, 0.00778, 0.05195,
              0.05103, 0.032, 0.01765, 0.01154, -0.00028, 0.01627, -0.00051,
              0.01439, -0.00937, -0.01135, 0.03692, 0.14959, 0.00803, 0.02424,
              -0.03548, 3.65223, -0.02515, -4.74563, 0.26259, -6.90124, 0.01583)

# fp <- attr(bw$Urban, "specfp")
# fp <- prepare_rhybrid(fp, rw_start = 2005, rw_dk = 1)

## Set some flags that are set in fitmod(), (later improve this code...)
# fp$ancsitedata <- TRUE
# fp$ancrt <- "both"
# fp$logitiota <- TRUE
# fp$rw_start <- 2005
# fp$incidmod <- "eppspectrum"

# param <- fnCreateParam(theta_ur, fp)
# fp_par <- update(fp, list = param)
```

Simulate the model.