---
title: "Basic example of sexual debut simulation"
author: "Kinh Nguyen"
date: "`r Sys.Date()`"
output: 
  html_vignette: 
    toc: true
    fig_width: 7
    fig_height: 5
bibliography: ref.bib
vignette: >
  %\VignetteIndexEntry{Basic example of sexual debut simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  comment = "#>"
)
```

This vignette develops the workflow for simulating the EPP-ASM model with sexual
debut. 

Sexual debut model has the following extra movements

- Base population in the sexual debut age range is inititated as inactive and
  then move to active based on the debut probability.
- Entrant population enters the model as sexual inactive and then moves to
  sexual active with the debut probability.
- There are separete HIV and ART pop for sexual inactive group, these also move
  from inactive to active.
- HIV and ART pop is splitted to single year in the debut age range, the older
  age-groups (max debut age + 1) are grouped as the original age-group but with
  new group indices.
- Incidence calculation ignores the inactive population (including those in HIV
  positive with and without treatment).
- Core population, HIV population, and ART population are assigned as their own
  class (`refClasses.R`).

```{r echo=FALSE, message=FALSE, load}
## checkout `debut-withClass` branch
library(magrittr)
devtools::load_all()
```

## Preparing EPP-ASM inputs

```{r, eppasm_inputs}
# Use Botswana for convinient
pjnz <- system.file("extdata/testpjnz", "Botswana2018.PJNZ", package="eppasm")
bw <- prepare_spec_fit(pjnz, proj.end=2021.5)
```

### EPP model with single set of parameter inputs

Prepare model parameters from parameter vector input
```{r, single_param}
theta_ur <- c(-0.63758, -2.76655, -1.26204, 1996.65945, 0.00778, 0.05195,
              0.05103, 0.032, 0.01765, 0.01154, -0.00028, 0.01627, -0.00051,
              0.01439, -0.00937, -0.01135, 0.03692, 0.14959, 0.00803, 0.02424,
              -0.03548, 3.65223, -0.02515, -4.74563, 0.26259, -6.90124, 0.01583)

fp <- attr(bw$Urban, "specfp")
# Constructing `fp$rt`
fp <- prepare_rhybrid(fp, rw_start = 2005, rw_dk = 1)

## Set some flags that are set in fitmod(), (later improve this code...)
fp$ancsitedata <- TRUE
fp$ancrt <- "both"
fp$logitiota <- TRUE
fp$rw_start <- 2005
fp$incidmod <- "eppspectrum"

# create rvec, iota, update frr,...
param <- fnCreateParam(theta_ur, fp)
fp_par <- update(fp, list = param)
```

Blindly fix some wrong inputs

```{r}
# Target pop < 0
negid <- fp_par$targetpop %>% {which(. < 0, TRUE)} %T>% {cat(fp_par$targetpop[.])}
fp_par$targetpop[negid] %<>% '*'(-1)
```
Simulate the model

### C++ version of the current model
```{r, simulate_model}
mod <- simmod(fp_par)
```

### R version of the current model
```{r}
modR <- simmod(fp_par, VERSION = "R")

expect_equal(prev(mod), prev(modR))
expect_equal(incid(mod), incid(modR))
```

## Demographic model (R only)

```{r, demographic, fig.height=6, fig.cap = "Figure: Demographic only"}
demR <- simmod(fp_par, VERSION = "R", 0)
plot(demR)
```

## Sexual debut (R only)

How to run:

- update fp with debut probability matrix (add parameters and indices):
  - if data by age-group (`single_year=FALSE`, default):
    - 4 columns by order: male, female, lower age, upper age
  - if data by single-age (`single_year=TRUE`):
    - 4 columns by order: male, female, age male, age female
- specify `VERSION="R"` and `.MODEL=2`

### Test

That the model with 100% debut probability should be similar with model without
debut:

```{r}
# All debut
db1 <- eppasm:::debut_pr
db1[, 1:2] <- 1

fp_db1 <- update_fp_debut(fp_par, db1)
modUni <- simmod(fp_db1, "R", 2)
```

Not exactly, only visually

```{r, fig.show='hold', fig.width=3, fig.height=3}
plot(prev(mod), prev(modUni)); abline(0,1)
plot(incid(mod), incid(modUni)); abline(0,1)
```

## Simulation scenarios

### By age-group [@Johnson:2009dd]

```{r echo = FALSE, results = 'asis'}
knitr::kable(eppasm:::debut_pr, caption="Table: Name")
```

```{r}
fp_js <- update_fp_debut(fp_par, eppasm:::debut_pr)
modjs <- simmod(fp_js, VERSION = "R", 2)
``` 

### By single-year pr up to 17 [@CavazosRehg:2009dl, Tab. 2, African-American]


```{r echo=FALSE}
# Males
male <- matrix(c(12, 0.85, 13, 0.72, 14, 0.58, 15, 0.42, 16, 0.28, 17, 0.18),
               ncol=2, byrow=TRUE)
# Females               
female <- matrix(c(12, 0.97, 13, 0.92, 14, 0.83, 15, 0.66, 16, 0.45, 17, 0.26),
                 ncol=2, byrow=TRUE)
debut_cv <- cbind(male[,2], female[,2], male[,1], female[,1])
debut_cv[, 1:2] <- 1 - debut_cv[, 1:2]
```

Assume all debut at 18

```{r}
debut_cv <- rbind(debut_cv, c(1, 1, 18, 18))
```

```{r echo = FALSE, results = 'asis'}
knitr::kable(debut_cv, caption="Table: @CavazosRehg:2009dl")
```

```{r}
fp_db_cv <- update_fp_debut(fp_par, debut_cv, NULL, single_year=TRUE)
modcv <- simmod(fp_db_cv, "R", 2)
```

### DHS data (Zimbabwe as an example)

Smooth and step-wise

```{r message=FALSE}
# devtools::install_github("ropensci/rdhs")
library(rdhs)

indicators <- dhs_indicators()
first_sex  <- grep('first sex', indicators[,"Definition"])
first_sex  <- first_sex[1:10]
myid       <- indicators[first_sex, "IndicatorId"]

ccc <- dhs_countries(returnFields=c("CountryName","DHS_CountryCode"))
zcc <- ccc[grep("Zimbabwe", ccc[, 2]), 1]

fsex <- dhs_data(countryIds = zcc, indicatorIds = myid, breakdown="all") %$%
        cbind.data.frame(perc=Value, agr=CharacteristicLabel, year=SurveyYear,
                         id=IndicatorId) %>% 
        .[-grep("Total", .[, "agr"]), ] %>% 
        subset(year==2015)

fsex$sex <- substring(fsex$id, nchar(as.character(fsex$id))-4,
                      nchar(as.character(fsex$id))-4) %>% factor %>% as.numeric
fsex$age <- substring(fsex$id, nchar(as.character(fsex$id))-1,
                          nchar(as.character(fsex$id))) %>% as.numeric

M_sm <- fsex %>% subset(sex==1) %$% smooth.spline(age, perc) %>% predict(15:30)
W_sm <- fsex %>% subset(sex==2) %$% smooth.spline(age, perc) %>% predict(15:30)
M_cs <- fsex %>% subset(sex==1) %$% approx(age, perc, 15:30, "constant")
W_cs <- fsex %>% subset(sex==2) %$% approx(age, perc, 15:30, "constant")

# Extrapolate to 30 when all are sexual experienced (debut probability = 1 in 30)
M_sm$y[M_sm$y > 100] <- 100
W_sm$y[W_sm$y > 100] <- 100
M_cs$y[is.na(M_cs$y)] <- max(M_cs$y, na.rm=TRUE)
W_cs$y[is.na(W_cs$y)] <- max(W_cs$y, na.rm=TRUE)
W_cs$y[length(W_cs$y)] <- 100
M_cs$y[length(M_cs$y)] <- 100
```

```{r echo=FALSE, fig.cap = "Figure: Probability of sexual debut", fig.height=6}
put(1,1, c(5,4,1,1))
fsex %$% pointplot(c(age, 26:30), c(perc, rep(NA, 5)), col=as.numeric(sex)+3,
                   xlab="Age", ylab="Percent", ylim=c(0,100))
lines(W_cs, col=5, lty=3); lines(W_sm, col=5)
lines(M_cs, col=4, lty=3); lines(M_sm, col=4)
legend("topleft", col=1:2+3, legend=c("M", "F"), lty=1)
```

```{r echo=FALSE}
zim_sm <- cbind.data.frame(data.frame(M_sm), data.frame(W_sm))[, c(2,4,1,3)]
zim_cs <- cbind.data.frame(data.frame(M_cs), data.frame(W_cs))[, c(2,4,1,3)]
zim_sm[, 1:2] <- zim_sm[, 1:2]/100
zim_cs[, 1:2] <- zim_cs[, 1:2]/100
```

Update parameters and run

```{r}
# Smooth data
fp_sm  <- update_fp_debut(fp_par, zim_sm, single_year=TRUE)
mod_sm <- simmod(fp_sm, "R", 2)
# Step-wise data
fp_cs <- update_fp_debut(fp_par, zim_cs, single_year=TRUE)
mod_cs <- simmod(fp_cs, "R", 2)
```

## Outputs

- `output$data` is the same as the old `output`
- all attributes are stored as fields and accessed directly by `$`, with the same names.

### Prevalence by models

```{r echo=FALSE, fig.cap = "Figure: Prevalence 15-49"}
put(1,1, c(5,4,1,1))
ylim <- c(0, max(prev(mod), prev(modcv), prev(modjs), prev(mod_sm), prev(mod_cs)))
lineplot(prev(mod), col=1, lwd=2, ylim=ylim)
lines(prev(modcv), col=2, lwd=2)
lines(prev(modjs), col=3, lwd=2)
lines(prev(mod_sm), col=4, lwd=2)
lines(prev(mod_cs), col=5, lwd=2)
legend("topleft", col=1:5, lwd=2,
       legend=c("EPP", "Cavazos-Rehg", "Johnson", "Zimbabwe smooth", "Zimbabwe step-wise"))
```

### Prevalence by ages

```{r echo=FALSE, fig.cap = "Figure: Prevalence by ages (Zimbabwe smooth)"}
put(1,1, c(5,4,1,1))
plot_prev(mod_sm, main="with debut")
```

Prevalence among sexually active population vs. total population and % sexually
active by age

```{r echo=FALSE, fig.cap = "Figure: Prevalence by ages (Zimbabwe smooth)"}
plot_prev(mod_sm, debut=TRUE, years=c(1990, 2000, 2010, 2018), main="with debut")
plot_prev(mod_sm, debut=TRUE, years=c(2000:2010), main="with debut")
```

### Prevalence by age over time

```{r echo=FALSE, fig.cap = "Figure: Prevalence by ages over time"}
plot_prev(mod_cs, stackbar=TRUE, separate=TRUE)
```

## References
