---
title: "Basic example stepping through EPP-ASM"
author: ""
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  comment = "#>"
)
```

This vignette develops the workflow for fitting the EPP-ASM model and generating outputs to provide as Spectrum inputs.

```{r, load}
#eppasm version >= 0.5.2
install.packages("devtool")
devtools::install_github("mrc-ide/eppasm@new-master")
library(eppasm)
devtools::load_all()
devtools::load_all(path = "~/Documents/eppasm/")
```

## Preparing EPP-ASM inputs

```{r, eppasm_inputs}
pjnz <- system.file("extdata/testpjnz", "Botswana2018.PJNZ", package="eppasm")
bw <- prepare_spec_fit(pjnz, proj.end=2021.5)
```

## Simulate model with direct incidence input (a la Spectrum

```{r, direct_incid}
fp <- prepare_directincid(pjnz)
mod <- simmod(fp)

prev(mod)
incid(mod)
```

## EPP model with single set of parameter inputs

Prepare model parameters from parameter vector input
```{r, single_param}
theta_ur <- c(-0.63758, -2.76655, -1.26204, 1996.65945, 0.00778, 0.05195,
              0.05103, 0.032, 0.01765, 0.01154, -0.00028, 0.01627, -0.00051,
              0.01439, -0.00937, -0.01135, 0.03692, 0.14959, 0.00803, 0.02424,
              -0.03548, 3.65223, -0.02515, -4.74563, 0.26259, -6.90124, 0.01583)

fp <- attr(bw$Urban, "specfp")
fp <- prepare_rhybrid(fp, rw_start = 2005, rw_dk = 1)

## Set some flags that are set in fitmod(), (later improve this code...)
fp$ancsitedata <- TRUE
fp$ancrt <- "both"
fp$logitiota <- TRUE
fp$rw_start <- 2005
fp$incidmod <- "eppspectrum"

fp <- prepare_anc_model(fp, attr(bw$Urban, "eppd"))

param <- fnCreateParam(theta_ur, fp)
fp_par <- update(fp, list = param)
```

Simulate the model once.

```{r, simulate_model}
mod <- simmod(fp_par)

prev(mod)
incid(mod)

## R version of the model
modR <- simmod(fp_par, VERSION = "R")

prev(modR)
incid(modR)
```


Prepare likelihood and calculate the likelihood once

```{r, calc_ll}
## Prior 
lprior(theta_ur, fp)

## Prepare likelihood data
likdat <- prepare_likdat(attr(bw$Urban, "eppd"), fp)

## Calculate likelihood
ll(theta_ur, fp, likdat)

## Components of likelihood calculation
ll_hhsage(mod, likdat$hhs.dat)

ll_ancsite(mod, fp_par,
           coef = c(fp_par$ancbias, fp_par$ancrtsite.beta),
           vinfl = fp_par$v.infl,
           dat = likdat$ancsite.dat)

ll_ancrtcens(mod, likdat$ancrtcens.dat, fp_par)
```


## Fitting the EPP-ASM model

```{r, fit_eppasm, cache=TRUE}
bwfit <- list()

bwfit$Urban <- fitmod(bw$Urban, eppmod = "rhybrid", rw_start = 2005,
                      B0=1e3, B=1e2, opt_iter = 1, number_k=50, optfit = TRUE)
bwfit$Rural <- fitmod(bw$Rural, eppmod = "rhybrid", rw_start = 2005,
                      B0=1e3, B=1e2, opt_iter = 1, number_k=50, optfit = TRUE)
```


When fitting, the random-walk based models only simulate through the end of the
data period. The `extend_projection()` function extends the random walk for $r(t)$
through the end of the projection period.

```{r, extend_projection}
bwfit <- lapply(bwfit, extend_projection, proj_years = 52)
```

## Simulating model outptus

```{r, tidy_output}

bwout <- Map(tidy_output, bwfit, "r-hybrid", "Botswana", names(bwfit))

```

## Pooling EPP subpopulation results

The function `aggr_specfit()` 
This involves simulating the model for all resamples in each subregion and summing the following `pop`, `hivpop`, and `artpop` arrays for each of the 3000 resamples to generate 3000 national outputs.

```{r, aggr_specfit}
bwaggr <- aggr_specfit(bwfit)
```
