---
title: "EPP-ASM to Spectrum inputs workflow"
author: "Jeff Eaton, Rob Glaubius, Tim Brown, John Stover"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  comment = "#>"
)
```

This vignette develops the workflow for fitting the EPP-ASM model and generating outputs to provide as Spectrum inputs.

```{r, load}
## eppasm version >= 0.5.2
## devtool::install_github("mrc-ide/eppasm@dev")
## library(eppasm)
devtools::load_all()
```

## Preparing EPP-ASM inputs

```{r, eppasm_inputs}
pjnz <- system.file("extdata/testpjnz", "Botswana2018.PJNZ", package="eppasm")
bw <- prepare_spec_fit(pjnz, proj.end=2021.5)
```

## Fitting the EPP-ASM model

```{r, fit_eppasm, cache=TRUE}
bwfit <- list()

bwfit$Urban <- fitmod(bw$Urban,
                      eppmod = "rhybrid", rw_start = 2005, fitincrr = FALSE,
                      B0=1e4, B=1e3, opt_iter = 1:2*5, number_k=50)
bwfit$Rural <- fitmod(bw$Rural,
                      eppmod = "rhybrid", rw_start = 2005, fitincrr = FALSE,
                      B0=1e4, B=1e3, opt_iter = 1:2*5, number_k=50)
```

When fitting, the random-walk based models only simulate through the end of the
data period. The `extend_projection()` function extends the random walk for $r(t)$
through the end of the projection period.

```{r, extend_projection}
bwfit <- lapply(bwfit[1], extend_projection, proj_years = 52)
```

## Pooling EPP subpopulation results

The function `aggr_specfit()` 
This involves simulating the model for all resamples in each subregion and summing the following `pop`, `hivpop`, and `artpop` arrays for each of the 3000 resamples to generate 3000 national outputs.

```{r, aggr_specfit}
bwaggr <- aggr_specfit(bwfit)

```

## Generating outputs 

Now generate outputs for the following by age, sex, and year:

* Total population
* HIV positive population
* New infections
* Non-HIV deaths
* HIV deaths

This is all a bit manual right now. In due course, will write better functions to assist with this.


```{r, generate_outputs}
library(magrittr)
library(data.table)

dimnm <- list(age = 15:80, sex = c("male", "female"), year = 1970:2021)

totpop <- lapply(bwaggr, apply, c(1, 2, 4), sum) %>%
  abind::abind(., rev.along=0, use.dnns=TRUE,
               new.names = c(dimnm, list(sampleid = seq_along(.)))) %>%
  melt %>%
  data.table(., outcome = "totpop")

hivpop <- lapply(bwaggr, function(x) x[ , , 2, ]) %>%
  abind::abind(., rev.along=0, use.dnns=TRUE,
               new.names = c(dimnm, list(sampleid = seq_along(.)))) %>%
  melt %>%
  data.table(., outcome = "hivpop")

infections <- lapply(bwaggr, attr, "infections") %>%
  abind::abind(., rev.along=0, use.dnns=TRUE,
               new.names = c(dimnm, list(sampleid = seq_along(.)))) %>%
  melt %>%
  data.table(., outcome = "infections")

natdeaths <- lapply(bwaggr, attr, "natdeaths") %>%
  abind::abind(., rev.along=0, use.dnns=TRUE,
               new.names = c(dimnm, list(sampleid = seq_along(.)))) %>%
  melt %>%
  data.table(., outcome = "natdeaths")

hivdeaths <- lapply(bwaggr, attr, "hivdeaths") %>%
  abind::abind(., rev.along=0, use.dnns=TRUE,
               new.names = c(dimnm, list(sampleid = seq_along(.)))) %>%
  melt %>%
  data.table(., outcome = "hivdeaths")

outputs_age <- rbind(totpop, hivpop, infections, natdeaths, hivdeaths)

outputs_age$outcome <- factor(outputs_age$outcome, c("totpop", "hivpop", "infections", "natdeaths", "hivdeaths"))

dcast(outputs_age[sampleid == 1], year + sex + age ~ outcome) %>%
  write.csv("bwaggr_ageoutputs_single-resample_v2.csv", row.names=FALSE)

dcast(outputs_age[ , .(value = median(value)), .(year, sex, age, outcome)],
      year + sex + age ~ outcome) %>%
  write.csv("bwaggr_ageoutputs_median_v2.csv", row.names=FALSE)

dcast(outputs_age[ , .(value = mean(value)), .(year, sex, age, outcome)],
      year + sex + age ~ outcome) %>%
  write.csv("bwaggr_ageoutputs_mean_v2.csv", row.names=FALSE)

```

For comparisons, the following annual aggregate outputs are generated:

* HIV prevalence among age 15-49
* HIV incidence rate among age 15-49
* ART coverage among age 15+

```{r, aggr_outputs}

hivprev <- sapply(bwaggr, prev) %>%
  "dimnames<-"(list(year = 1970:2021, sampleid = seq_len(ncol(.)))) %>%
  melt %>%
  data.table(., outcome = "prev15to49")

hivincid <- sapply(bwaggr, incid) %>%
  "dimnames<-"(list(year = 1970:2021, sampleid = seq_len(ncol(.)))) %>%
  melt %>%
  data.table(., outcome = "incid15to49")

artcov <- sapply(bwaggr, artcov15plus) %>%
  "dimnames<-"(list(year = 1970:2021, sampleid = seq_len(ncol(.)))) %>%
  melt %>%
  data.table(., outcome = "artcov15plus")

outputs_aggr <- rbind(hivprev, hivincid, artcov)

outputs_aggr$outcome <- factor(outputs_aggr$outcome,
                               c("prev15to49", "incid15to49", "artcov15plus"))

dcast(outputs_aggr[sampleid == 1], year ~ outcome) %>%
  write.csv("bwaggr_aggroutputs_single-resample_v2.csv", row.names=FALSE)

dcast(outputs_aggr[ , .(value = median(value)), .(year, outcome)],
      year ~ outcome) %>%
  write.csv("bwaggr_aggroutputs_median_v2.csv", row.names=FALSE)

dcast(outputs_aggr[ , .(value = mean(value)), .(year, outcome)],
      year ~ outcome) %>%
  write.csv("bwaggr_aggroutputs_mean_v2.csv", row.names=FALSE)

```

## Parameter vectors for single resample results

Storing the paremter vectors for the single resample outputs so that we can 
reproduce and step through the simulations if necessary.

```{r, theta_single}
theta_urban1 <- dput(c(bwfit$Urban$resample[1,]))

theta_rural1 <- dput(c(bwfit$Rural$resample[1,]))
```	


# National EPP fit outputs

Reproduce the above with a national EPP fit rather than Urban / Rural fits

```{r national}

devtools::load_all("~/Documents/Code/R/eppasmdata")

bw <- prepare_national_fit(pjnz, proj.end=2021.5)
attr(bw, "eppd") <- create_eppd("Botswana")

bwfit <- fitmod(bw,
                eppmod = "rhybrid", rw_start = 2005, fitincrr = "linincrr",
                B0=1e4, B=1e3, opt_iter = 1:2*5, number_k=50)
bwfit <- extend_projection(bwfit, 52)

bwfit$param <- lapply(seq_len(nrow(bwfit$resample)), function(ii) fnCreateParam(bwfit$resample[ii,], bwfit$fp))
fplist <- lapply(bwfit$param, function(par) update(bwfit$fp, list=par))
bwnat <- lapply(fplist, simmod)

bwnat <- sim_mod_list(bwfit)


dimnm <- list(age = 15:80, sex = c("male", "female"), year = 1970:2021)

totpop <- lapply(bwnat, apply, c(1, 2, 4), sum) %>%
  abind::abind(., rev.along=0, use.dnns=TRUE,
               new.names = c(dimnm, list(sampleid = seq_along(.)))) %>%
  melt %>%
  data.table(., outcome = "totpop")

hivpop <- lapply(bwnat, function(x) x[ , , 2, ]) %>%
  abind::abind(., rev.along=0, use.dnns=TRUE,
               new.names = c(dimnm, list(sampleid = seq_along(.)))) %>%
  melt %>%
  data.table(., outcome = "hivpop")

infections <- lapply(bwnat, attr, "infections") %>%
  abind::abind(., rev.along=0, use.dnns=TRUE,
               new.names = c(dimnm, list(sampleid = seq_along(.)))) %>%
  melt %>%
  data.table(., outcome = "infections")

natdeaths <- lapply(bwnat, attr, "natdeaths") %>%
  abind::abind(., rev.along=0, use.dnns=TRUE,
               new.names = c(dimnm, list(sampleid = seq_along(.)))) %>%
  melt %>%
  data.table(., outcome = "natdeaths")

hivdeaths <- lapply(bwnat, attr, "hivdeaths") %>%
  abind::abind(., rev.along=0, use.dnns=TRUE,
               new.names = c(dimnm, list(sampleid = seq_along(.)))) %>%
  melt %>%
  data.table(., outcome = "hivdeaths")

outputs_age <- rbind(totpop, hivpop, infections, natdeaths, hivdeaths)

outputs_age$outcome <- factor(outputs_age$outcome, c("totpop", "hivpop", "infections", "natdeaths", "hivdeaths"))

dcast(outputs_age[sampleid == 1], year + sex + age ~ outcome) %>%
  write.csv("bwnat_ageoutputs_single-resample_v2.csv", row.names=FALSE)

dcast(outputs_age[ , .(value = median(value)), .(year, sex, age, outcome)],
      year + sex + age ~ outcome) %>%
  write.csv("bwnat_ageoutputs_median_v2.csv", row.names=FALSE)

dcast(outputs_age[ , .(value = mean(value)), .(year, sex, age, outcome)],
      year + sex + age ~ outcome) %>%
  write.csv("bwnat_ageoutputs_mean_v2.csv", row.names=FALSE)



hivprev <- sapply(bwnat, prev) %>%
  "dimnames<-"(list(year = 1970:2021, sampleid = seq_len(ncol(.)))) %>%
  melt %>%
  data.table(., outcome = "prev15to49")

hivincid <- sapply(bwnat, incid) %>%
  "dimnames<-"(list(year = 1970:2021, sampleid = seq_len(ncol(.)))) %>%
  melt %>%
  data.table(., outcome = "incid15to49")

artcov <- sapply(bwnat, artcov15plus) %>%
  "dimnames<-"(list(year = 1970:2021, sampleid = seq_len(ncol(.)))) %>%
  melt %>%
  data.table(., outcome = "artcov15plus")

outputs_aggr <- rbind(hivprev, hivincid, artcov)

outputs_aggr$outcome <- factor(outputs_aggr$outcome,
                               c("prev15to49", "incid15to49", "artcov15plus"))

dcast(outputs_aggr[sampleid == 1], year ~ outcome) %>%
  write.csv("bwnat_aggroutputs_single-resample_v2.csv", row.names=FALSE)

dcast(outputs_aggr[ , .(value = median(value)), .(year, outcome)],
      year ~ outcome) %>%
  write.csv("bwnat_aggroutputs_median_v2.csv", row.names=FALSE)

dcast(outputs_aggr[ , .(value = mean(value)), .(year, outcome)],
      year ~ outcome) %>%
  write.csv("bwnat_aggroutputs_mean_v2.csv", row.names=FALSE)
